---
title: "Drug CV: `r (params$x)`"
author: "CAMARADES / AD-SOLES-CT"
date: "`r format(Sys.time(), '%d %B %Y')`"

output: 
  tint::tintPdf:
  template: "tintPdf-template.tex"
  pdf_engine: lualatex


header-includes:
- \usepackage{booktabs}
- \usepackage[utf8]{inputenc}
- \usepackage{amssymb}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
- \DeclareUnicodeCharacter{2009}{\,}
- \DeclareUnicodeCharacter{3B1}{$\alpha$}
- \DeclareUnicodeCharacter{2265}{$\geq$}
- \DeclareUnicodeCharacter{2261}{$\equiv$}
# - \DeclareUnicodeCharacter{FF08}{\(}
# - \DeclareUnicodeCharacter{FF09}{\)}



tables: yes

params:
  x: cilostazol
---

```{r library, echo=FALSE, include=FALSE}
knitr::opts_chunk$set(tidy = FALSE, echo=FALSE, warning=FALSE, cache.extra = packageVersion('tint'))

options(htmltools.dir.version = FALSE)
library(googlesheets4)
library(dplyr)
library(knitr)
library(tint)
library(tidyr)
library(kableExtra)
library(stringr)
library(ggplot2)
library(RColorBrewer)
library(ggrepel)
library(textclean)
library(aws.s3)
library(magick)
library(stringr)
library(htmltools)
library(httr)
library(xml2)
library(rvest)
library(meta)
library(metafor)
library(metaviz)
library(jsonlite)
library(DBI)
```


```{r setup, include=FALSE}
source("configure.R")
source("getClinicalTrialsData.R")

drugSummary<-googlesheets4::read_sheet(googleSheetId, sheet = "drugSummary")

publicationList <-googlesheets4::read_sheet(googleSheetId, sheet = "publicationList") 
clinicalStudyList <- RMySQL::dbReadTable(con, "ReLiSyRClinicalStudies")
invivoStudies<- RMySQL::dbReadTable(con, "ad_invivo_citations")
```

# ReLiSyR data
## Overview
```{r drugscores, echo=FALSE, warning=F}
selecteddrugranklist <- drugSummary %>%
  filter(Drug %in% params$x)%>%
  select(
    "Drug",
    "drugScore",
    "nPublication",
    "efficacyScore"  ,
    "safetyScore"  ,
    "nParticipants" ,
    "qualityScore")%>%t()

row.names(selecteddrugranklist)<- c("Drug","Clinical Drug Score", "Number of clinical publications","Median Efficacy Score (-2 - 4)","Median Safety Score (0-3)","Median number of participants","Median Quality Score (out of 24)")


if(ncol(selecteddrugranklist)>0){
kbl(selecteddrugranklist, digits=3)%>%
   row_spec(0, bold=TRUE)} else print("No scored ReLiSyR clinical publications for selected drug")



```

```{r fig-nocap-margin-first, fig.margin=TRUE, fig.width=4, fig.height=4, fig.cap="Violin plot of drug scores, selected drug shown in red", cache=TRUE, echo=FALSE, warning=FALSE}
drugScores<-data.frame(x=drugSummary$Drug,y=drugSummary$drugScore)
selectedDrugScore<-drugScores%>%
  filter(drugScores$x==params$x)%>%
  select("y")%>%
  as.numeric()

if(!is.na(selectedDrugScore)){
drugScoresPlot<-ggplot(drugScores, aes(y=y, x=""))+
  labs(y="Drug Score", x="")+
  geom_violin(trim=F, fill="cadetblue2")+
  geom_dotplot(dotsize=0.8, binaxis='y', stackdir='center', binwidth=0.25)+
  annotate(geom="point", x="", y=selectedDrugScore, shape=13, colour="red3", size=5,stroke=2)

drugScoresPlot
}

```

```{r fig-nocap-margin-second, fig.margin=TRUE, fig.width=4, fig.height=4, cache=TRUE, echo=FALSE, fig.cap="Clinical subscores scatter plot, selected drug shown in black"}
if(!is.na(selectedDrugScore)){
drugbubble<-function(){
  df<-drugSummary[drugSummary$Drug==params$x,]
  drugSafety<-df$safetyScore
  drugEfficacy<-df$efficacyScore
  
  livedrugrankchart <- ggplot(
    drugSummary,
    aes(
      x = safetyScore,
      y = efficacyScore,
      size= nParticipants,
      color= qualityScore
    ))+ 
    scale_color_gradient(low="yellow", high="red", limits=c(0,24))+ 
    geom_point(alpha=1)+
    # scale_size("studySizeScore", range=c(1,4))
    theme(legend.position = "right")+
    labs(x="Safety Score",
         y="Efficacy Score",
         color="Quality Score",
         size="nParticipants")+
    xlim(0,3)+
    ylim(-2,4)
  #+lims(size=c(1,4))
  
  
  drugbubble<-livedrugrankchart+annotate(geom = "point", x =  drugSafety, y =  drugEfficacy, shape=13, size=5,stroke=1, color="black")
  
  drugbubble
  
}

drugbubble()
}
```
\ 

Current results summary of (i) clinical review: For each publication, we calculated a distance score based on Euclidean distance of efficacy [*E*] and safety [*S*] scores weighted by [*w*] comprising of quality score, *q* and study size, *nParticipants*:

$$\textrm{distance score} = \sqrt {|\overrightarrow{E}(wE)^2 + (wS)^2|} \times \overrightarrow{d}$$

where 

$$ w = \sqrt { (\frac{\sum q}{24})^2 + (\frac{nParticipants}{1000})^2}$$

$\overrightarrow{E}$ = direction of effect of efficacy, where:  
if $E$ < 0, $\overrightarrow{E}$ = -1; and if $E$ > 0, $\overrightarrow{E}$ = +1;  

and
$\overrightarrow{d}$ = direction of effect of distance score, where

$\overrightarrow{d} = +1$ if $(\overrightarrow{E}(wE)^2 + (wS)^2) > 0$;

$\overrightarrow{d} = -1$ if $(\overrightarrow{E}(wE)^2 + (wS)^2) < 0$  


Separately, we calculate median subscores for efficacy, safety, quality, and number of participants across all publications for each drug; (ii) *animal in vivo* and (iii) *in vitro* review are reported here using pooled effect sizes by random effects meta-analysis based on median survival (animal in vivo) and cell death (in vitro) for drugs where these are described in 5 or more experiments in 3 or more publications.

\newpage


## Clinical Review

Table of number of clinical studies and number of patients for `r params$x` by disease, study type and phase. 

```{r study design overview, warning=F,message=F}
clinicalPubs <- publicationList %>%
  filter(Drug %in% params$x) %>%
  select(
    "Title",
    "Disease",
    "Year",
    "studyType",
    "phase",
    "nPatients"
  )

clinicalPubs$studyType[clinicalPubs$studyType=="Interventional trial"] = "interventional"
clinicalPubs$phase[clinicalPubs$phase=="not reported"] = "Not specified"
clinicalPubs$phase[clinicalPubs$phase== 'c("Phase II", "Phase III")'] = "Phase II/III"


studyTypeCount<-function(){
  
  studyTypeOverview<-clinicalPubs%>%
    group_by(Disease, studyType)%>%
    summarise(nStudies = n())
  
  phaseOverview<-clinicalPubs%>%
    filter(clinicalPubs$studyType=="interventional")%>%
    group_by(Disease, phase)%>%
    summarise(nStudies=n())%>%
    mutate(studyType="interventional")
  
  observational<-studyTypeOverview%>%
    filter(studyType =="observational")%>%
    mutate(phase="NA")
  
  studyOverview<-rbind(phaseOverview,observational)
  studyOverview<-studyOverview[,c(1,4,2,3)]
  return(studyOverview)
}

studySizeOverview<-function(){
  studySize<-clinicalPubs%>%
    group_by(Disease,studyType)%>%
    summarise(nPatients=sum(nPatients))
  
  studySizePhase<-clinicalPubs%>%
    filter(clinicalPubs$studyType=="interventional")%>%
    group_by(Disease,phase)%>%
    summarise(nPatients=sum(nPatients))%>%
    mutate(studyType="interventional")
  
  observational<-studySize%>%
    filter(studyType =="observational")%>%
    mutate(phase="NA")
  studySizeOverview<-rbind(studySizePhase, observational)
  studySizeOverview<-studySizeOverview[,c(3,1,2)]
  return(studySizeOverview)
  
}


clinicalOverview<-left_join(studyTypeCount(),studySizeOverview())%>%as.data.frame()

func <- function(z) if (is.numeric(z)) sum(z) else '' 
sumrow <- as.data.frame(lapply(clinicalOverview, func))
sumrow[1] <- "Total"
clinicalOverview <- rbind(clinicalOverview, sumrow)


colnames(clinicalOverview)<-c("Disease", "Study Type", "Phase", "Number of Studies", "Number of Patients")
if(nrow(clinicalOverview)>1){
kbl(clinicalOverview) %>%
  row_spec(0, bold=TRUE)%>%
  row_spec((dim(clinicalOverview)-1)[1], hline_after = TRUE)%>%

  row_spec(dim(clinicalOverview)[1], bold = T,
           # background = "#ECECEC", 
           hline_after = TRUE)%>%
  collapse_rows(columns = 1)
}

```




## Clinical Publications


```{r clinical publications}

clinicalPubs1<-publicationList%>%
  filter(Drug %in% params$x) %>%
  select(
    "StudyIdStr",
    "Title",
    "Disease",
    "Year",
    "efficacyScores",
    "safteyScores",
    "nPatients",
    "qualityScore",
    "distanceScore",
    "PdfRelativePath"
  )

if(nrow(clinicalPubs1)>0){

urls<-clinicalPubs1$PdfRelativePath

#clinicalPubs1$Title<-substr(clinicalPubs1$Title, start=1, stop=50)




clinicalPubs2<-clinicalPubs1%>%
  select(
    "Title",
    "Disease",
    "Year",
    "efficacyScores",
    "safteyScores",
    "nPatients",
    "qualityScore",
    "distanceScore")

clinicalPubs2$Title<-replace_non_ascii(clinicalPubs2$Title)

colnames(clinicalPubs2)<-c("Title","Disease","Year","[E]","[S]","nParticipants","[Q]","Distance score")

x<- kbl(clinicalPubs2, format="latex", digits=3, longtable=TRUE)%>%
  column_spec( 1, link = urls, width="7cm")%>%
  row_spec( 0, bold=TRUE)

x}
#%>%
#  column_spec(1, width = "7cm")



```

## Included ReLiSyR clinical publications for `r params$x` that have yet to be annotated

```{r included clinical papers}


includedNonAnnotatedClinicalStudyList <-clinicalStudyList%>%
  filter(tolower(Drug) %in% tolower(params$x))%>%
  filter(!idStr %in% clinicalPubs1$StudyIdStr)%>%
    select(Title,
         Author,
         Year, 
         Disease, 
         Link
         )

includedNonAnnotatedClinicalStudyListLinks <- includedNonAnnotatedClinicalStudyList$Link

if(nrow(includedNonAnnotatedClinicalStudyList)>0){
kbl(includedNonAnnotatedClinicalStudyList[, 1:4], longtable = TRUE)%>%
  column_spec(1, link = includedNonAnnotatedClinicalStudyListLinks, width = "8cm")%>%
  column_spec(2:4, width = "2cm")
} else print("No included studies for selected drug in ReLiSyR clinical review")

```

## Animal in vivo papers
Publications for `r params$x` in Alzheimer's disease experimental models from the [Alzheimer's Disease Systematic Online Living Evidence Summary (AD-SOLES)]("https:/camarades.shinyapps.io/ad-soles")


```{r animal in vivo papers}

selecteddruganimalpubtable <- invivoStudies %>%
  filter(tolower(intervention) %in% params$x) %>%
  select(
    title, 
    author, 
    journal,
    year, 
    doi
  )%>%
  mutate(author = str_trunc(author, 30))




if(nrow(selecteddruganimalpubtable)>0){
  

selecteddruganimalpubtable <- selecteddruganimalpubtable%>%mutate(across(everything(), replace_non_ascii))
selecteddruganimalpublinks <- paste0("https://doi.org/" , selecteddruganimalpubtable$doi)

  kbl(selecteddruganimalpubtable, longtable = TRUE, row.names = FALSE)%>%
    column_spec(1, link = selecteddruganimalpublinks, width = "5cm")%>%
    column_spec(2:3, width = "2cm")%>%
    column_spec(4, width = "1cm")%>%
    column_spec(5, width = "4cm")
  
} else print("No animal in vivo studies identified for selected drug in AD-SOLES")
    
    
  
```


## Clinical trial registry data

Alzheimer disease or Alzheimer dementia studies from ClinicalTrials.gov for `r params$x`

```{r AD clinicaltrials, message = FALSE, warning = FALSE, results = "asis"}
  
ADClinicalTrials <- getADClinicalTrialsData(params$x)


if(!is.null(ADClinicalTrials)){
  ADClinicalTrials <- ADClinicalTrials %>%
    select(-studyType)%>%
    mutate(overallStatus = str_replace_all(tolower(overallStatus), "_", " "))%>%
    mutate(description = str_trunc(description, 500))%>%
    rename(`intervention description` = description)
  
  ADClinicalTrialsLinks <- paste0("https://clinicaltrials.gov.uk/study/",ADClinicalTrials$nctId)
  
  kbl(ADClinicalTrials, longtable = TRUE, row.names = FALSE)%>%
    column_spec(1, link = ADClinicalTrialsLinks, width = "2cm")%>%
    column_spec(2, width = "3cm")%>%
    column_spec(3, width = "1.7cm")%>%
    column_spec(4, width = "2.5cm")%>%
    column_spec(5, width = "1cm")%>%
    column_spec(6, width = "4cm")
  
} else print("No AD studies identified for selected drug from ClinicalTrials.gov")

```


Studies for conditions (neurodegenerative diseases, vascular dementia, cerebral small vessel disease, stroke) from ClinicalTrials.gov for `r params$x`

```{r other diseases clinicaltrials, message = FALSE, warning = FALSE, results = "asis"}
  
OtherClinicalTrials <- getOtherNeuroClinicalTrialsData(params$x)


if(!is.null(OtherClinicalTrials)){
  OtherClinicalTrials <- OtherClinicalTrials %>%
    select(-studyType)%>%
    mutate(overallStatus = str_replace_all(tolower(overallStatus), "_", " "))%>%
    mutate(description = iconv(description, to = "ASCII"))%>%
    mutate(description = str_trunc(description, 500))%>%
    rename(`intervention description` = description)
  

  
  OtherClinicalTrialsLinks <- paste0("https://clinicaltrials.gov.uk/study/",OtherClinicalTrials$nctId)
  

  kbl(OtherClinicalTrials, longtable = TRUE, row.names = FALSE)%>%
    column_spec(1, link = OtherClinicalTrialsLinks, width = "2cm")%>%
    column_spec(2, width = "3cm")%>%
    column_spec(3, width = "1.7cm")%>%
    column_spec(4, width = "2.5cm")%>%
    column_spec(5, width = "1cm")%>%
    column_spec(6, width = "4cm")
  
} else print("No studies identified for selected drug from ClinicalTrials.gov")

```


## ChEMBL database
[ChEMBL](https://www.ebi.ac.uk/chembl/) is a manually curated database of bioactive molecules with drug-like properties by [EMBL's European Bioinformatics Institute](https://www.ebi.ac.uk/).

```{r chembl data, warning = F, message = F, results = "asis"}

drug <- params$x
drug <- str_replace(drug, "\\s", "%20")

url <- paste0("https://www.ebi.ac.uk/chembl/api/data/molecule?pref_name__icontains=", drug)
data <- read_html(url)
molecules <- data %>%html_nodes("molecules")%>%html_children()

for(n in 1:length(molecules)){
  data<-molecules[n]
  name <- data %>%html_nodes("pref_name")%>%html_text()
  chemblIDs <- data %>%html_nodes("molecule_chembl_id")%>%html_text()%>%unique()
  synonyms <- data %>%html_nodes("molecule_synonym")%>%html_text()%>%unique()%>%paste(collapse="; ")
  # SMILES <- data %>% html_node("canonical_smiles")%>%html_text()  
  mol_formula<-data%>%html_node("full_molformula")%>%html_text()
  mol_weight<-data%>%html_node("full_mwt")%>%html_text()
  mol_inchi<- data%>%html_node("standard_inchi")%>%html_text()
  mol_type<-data%>%html_node("molecule_type")%>%html_text()
  natural_product <- data %>%html_node("natural_product")%>%html_text()
  natural_product <- ifelse(natural_product =="", FALSE, TRUE)
  oral <- data %>%html_node("oral")%>%html_text()
  oral <- ifelse(oral =="", FALSE, TRUE)
  class <- data%>%html_node("usan_stem_definition")%>%html_text()
  ro5Violations <- data%>%html_node("num_ro5_violations")%>%html_text()
  ro5 <- ifelse(ro5Violations == "", TRUE, FALSE)
  ro3_pass <- data %>%html_node("ro3_pass")%>%html_text()
  ro3 <- ifelse(ro3_pass == "Y", TRUE, FALSE)
  max_phase <- data %>% html_node("max_phase")%>%html_text()
  availability_type <- data %>%html_node("availability_type")%>%html_text()
  availability_valid_type<-c(0,1,2)
  type<-pmatch(availability_type, availability_valid_type)
  availability <- switch(type, 
                         availability = "discontinued",
                         availability = "prescription only",
                         availability = "over the counter"
  )
  
  
  outputi<- paste0("<h4>", name, "</h4>",
                   "<p>ChemBLID: ", chemblIDs, 
                   "</p><p>Synonyms: ", synonyms,
                   # "</p><p>SMILES: ", SMILES,
                   "</p><p>Molecule type: ", mol_type,
                   "</p><p>Molecule class: ", class,
                   "</p><p>Molecular formula: ", mol_formula,
                   "</p><p>Molecular weight: ", mol_weight,
                   "</p><p>Natural product: ", natural_product,
                   "</p><p>Oral formulation: ", oral, 
                   "</p><p>Rule of 5 met: ", ro5,
                   "</p><p>Rule of 3 met: ", ro3,
                   "</p><p>Max Clinical Trial Phase: ", max_phase,
                   "</p><p>Availability type: ", availability,
                   "</p><p></p><p></p>"
  )
  
  if(!exists("combinedOutput")) combinedOutput <- outputi else combinedOutput <- c(combinedOutput, outputi)
  
}
combinedOutput


```

